---
title: "Example analysis - Woodman et al 2019"
author: "Samuel Woodman"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example analysis - Woodman et al 2019}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


```{r setup, message=FALSE}
library(dplyr)
library(eSDM)
library(sf)

source(system.file("eSDM_vignette_helper.R", package = "eSDM"), local = TRUE, echo = FALSE)
```


## Overview
`eSDM` allows users to create ensembles of predictions from species distribution models (SDMs), either using the `eSDM` GUI or manually in R using `eSDM` functions. This vignette demonstrates creating and evaluating ensembles using `eSDM` functions by manually performing the example analysis from Woodman _et al._ (2019). The example analysis explores differences between blue whale SDM predictions for the California Current Ecosystem (CCE) from Becker _et al._ (2016; i.e., Model_B), Hazen _et al._ (2017; i.e., Model_H), and Redfern _et al._ (2017; i.e., Model_R). It also creates and evaluates ensembles of the predictions, with associated uncertainty. See Woodman _et al._ (2019) for additional details, and in particular Table 1 for details about differences between the models.

In this vignette, the three sets of predictions and the validation data are read from .rds files because the original files were too large to be included in the package. The sections where these data are imported includes the code for reading the data from their original files, although this code has been commented out. The original files can be downloaded through the GUI or at https://github.com/smwoodman/eSDM-data. 

This document contains code for plotting predictions. However, by default some of the plotting code is not run because it can take several minutes (these code chunks contain the comment "code not run"). If desired, you may run these code chunks manually in R. 

Before using data from this example analysis, please see the README.txt file for proper citation information (located at `system.file("extdata/README.txt", package = "eSDM"`) and contact the author.


## Import SDM predictions

The first step of the example analysis is to import and process the Model_B, Model_H, and Model_R predictions, along with their standard error (SE) values. Use `pts2poly_centroids` to create `sf` objects from polygon centroids from CSV files, `raster::raster` to import raster files, and `sf::st_read` to import GIS files. The dimensions of the Model_B and Model_H predictions are 0.09 x 0.09 degrees and 0.25 x 0.25 degrees, respectively; the second argument of `pts2poly_centroids` is half the length of one side of the polygon. GIS files already have a defined geometry that is read by `st_read`. Before overlaying predictions, you must ensure the following:

* All predictions are `sf` objects. See below for examples of converting a CSV file of grid centroids or a `raster` object to an `sf` object.
* The geometries of the predictions are valid; this can be checked using `sf::st_is_valid`.
* The geometries of the predictions have a defined coordinate reference system (crs); this can be checked using `sf::st_crs`.
* The predictions have the same longitudinal range (i.e., either [0, 360] or [-180, 180]); this can be checked using `sf::st_bbox`. You can use `sf::st_wrap_dateline` to convert an `sf` object to the longitudinal range [-180, 180], but note that this will cause plots to span [-180, 180] as well.

We can visualize the SDM predictions after importing and processing them. The plots below make up Fig. 3 in Woodman _et al._ (2019). This vignette uses the custom `plot_sf_3panel` function (in 'vignette_helper.R' located at `system.file("vignette_helper.R", package = "eSDM"`) for plotting. `plot_sf_3panel` and `tmap_sdm` (used later in this vignette) were not included as functions in the `eSDM` package because they are specific to the example analysis region and SDMs. However, they can provide guidelines and a framework for plotting SDMs using the `sf` and `tmap` packages, allowing you to adapt these functions to your specific plotting needs.

```{r}
# Import, process, and plot Model_B predictions
# model.b <- read.csv("Predictions_Beckeretal2016.csv")
model.b.sf <- readRDS(system.file("extdata/Predictions_Beckeretal2016.rds", package = "eSDM")) %>% 
  eSDM::pts2poly_centroids(0.09 / 2, crs = 4326) %>%
  st_wrap_dateline() %>%
  st_set_agr("constant")

model.b.sf

# Make base map
map.world <- eSDM::gshhg.l.L16

# Other option for making base map
# map.world <- st_geometry(st_as_sf(maps::map('world', plot = FALSE, fill = TRUE)))
```

```{r, fig.width=7, fig.height=3}
plot_sf_3panel(
  model.b.sf, "pred_bm", main.txt = "Model_B - ", map.base = map.world, 
  x.axis.at = c(-130, -125, -120)
)
```

```{r}
# Import, process, and plot Model_H predictions
# model.h <- read.csv("Predictions_Hazenetal2017.csv")
model.h.sf <- readRDS(system.file("extdata/Predictions_Hazenetal2017.rds", package = "eSDM")) %>% 
  dplyr::select(lon, lat, pred_bm, se) %>%
  eSDM::pts2poly_centroids(0.25 / 2, crs = 4326, agr = "constant")

model.h.sf
```

```{r, fig.width=7, fig.height=3}
plot_sf_3panel(
  model.h.sf, "pred_bm", main.txt = "Model_H - ", map.base = map.world, 
  x.axis.at = c(-135, -130, -125, -120)
)
```

```{r}
# Import, process, and plot Model_R predictions
# model.r <- st_read("Shapefiles/Predictions_Redfernetal2017.shp")
model.r.sf <- readRDS(system.file("extdata/Predictions_Redfernetal2017.rds", package = "eSDM")) %>% 
  st_make_valid() %>%
  st_set_agr("constant")

model.r.sf
```

```{r, fig.width=7, fig.height=3}
plot_sf_3panel(
  model.r.sf, "pred_bm", main.txt = "Model_R - ", map.base = map.world, 
  x.axis.at = c(-130, -125, -120)
)
```


```{r, eval=FALSE}
# Example code for converting raster to sf object; code not run
logo <- raster::raster(system.file("external/rlogo.grd", package="raster"))
logo.sf <- as(logo, "SpatialPolygonsDataFrame") %>% 
  sf::st_as_sf()
```


## Overlay predictions

Because the original predictions have both different spatial resolutions and coordinate systems, we must overlay them onto the same base geometry. See Woodman _et al._ (2019), the `eSDM` GUI manual, or the `overlay_sdm` documentation for details about the overlay process. For the example analysis, we use the geometry of Model_R as the base geometry. However, first we must import and process the study area and erasing polygons, with which we clip and erase land from the base geometry, respectively. We also visualize the base geometry.

```{r}
# Study area polygon
poly.study <- st_read(system.file("extdata/Shapefiles/Study_Area_CCE.shp", package = "eSDM")) %>%
  st_geometry() %>% 
  st_transform(st_crs(model.r.sf))

# Erasing polygon; clip to the buffered study area polygon reduces future computation time
poly.erase <- eSDM::gshhg.l.L16 %>%
  st_transform(st_crs(model.r.sf)) %>%
  st_make_valid() %>%
  st_crop(st_buffer(poly.study, 100000))

# Create the base geometry; st_erase() function defined in eSDM_vignette_helper.R
base.geom <- model.r.sf %>%
  st_geometry() %>%
  st_erase(poly.erase) %>% 
  st_intersection(poly.study) %>%
  st_cast("MULTIPOLYGON")
```

```{r, fig.width=5, fig.height=7}
# Visualize the base geometry
plot(st_transform(base.geom, 4326), col = NA, border = "black", axes = TRUE)
plot(map.world, add = TRUE, col = "tan", border = NA)
graphics::box()
```

Next, we convert any associated uncertainty values to variance. Uncertainty values must be overlaid as variance values to remain statistically valid. 

```{r}
# Convert SE values to variance
model.b.sf <- model.b.sf %>% 
  mutate(variance = se^2) %>% 
  dplyr::select(pred_bm, se, variance)
model.h.sf <- model.h.sf %>% 
  mutate(variance = se^2) %>% 
  dplyr::select(pred_bm, se, variance)
model.r.sf <- model.r.sf %>% 
  mutate(variance = se^2) %>% 
  dplyr::select(pred_bm, se, variance)
```

Now we can overlay the original predictions onto the base geometry. Note that because we clipped and erased land from the base geometry, we cannot simply use the original Model_R predictions and geometry. However, we can 'overlay' the Model_R predictions by matching indices of base geometry polygons and Model_R prediction values (`over3.sfb` in the code below).

```{r}
# Perform overlay, and convert overlaid uncertainty values to SEs
over1.sf <- eSDM::overlay_sdm(base.geom, st_transform(model.b.sf, st_crs(base.geom)), c("pred_bm", "variance"), 50) %>% 
  mutate(se = sqrt(variance))
over2.sf <- eSDM::overlay_sdm(base.geom, st_transform(model.h.sf, st_crs(base.geom)), c("pred_bm", "variance"), 50) %>% 
  mutate(se = sqrt(variance))
over3.sf <- eSDM::overlay_sdm(base.geom, model.r.sf, c("pred_bm", "variance"), 50) %>% 
  mutate(se = sqrt(variance))

over3.sfb <- model.r.sf %>% 
  st_set_geometry(NULL) %>% 
  dplyr::select(pred_bm, variance) %>% 
  st_sf(geometry = base.geom, agr = "constant") %>% 
  dplyr::mutate(se = sqrt(variance))
all.equal(over3.sf, over3.sfb)
rm(over3.sfb)

```

We also could plot the overlaid predictions to show that the overlaid distribution patterns are very similar to those of the original predictions.

```{r, fig.width=7, fig.height=3, eval=FALSE}
# Plot overlaid predictions; code not run
plot_sf_3panel(over1.sf, "pred_bm", main.txt = "Overlaid Model_B - ", map.base = map.world)
plot_sf_3panel(over2.sf, "pred_bm", main.txt = "Overlaid Model_H - ", map.base = map.world)
plot_sf_3panel(over3.sf, "pred_bm", main.txt = "Overlaid Model_R - ", map.base = map.world)
```


## Calculate evaluation metrics

To evaluate predictions and create ensembles with weights based on evaluation metrics, we must load and process the validation data sets for use with `evaluation_metrics`. The validation data must be points of class `sf` with a column with either binary presence/absence or count data. For the example analysis, we use three validation sets in the example analysis, line transect data (Becker _et al._ 2016), home range data (Irvine _et al._ 2014), and these two data sets combined. Because our validation data are binary (i.e., presence/absence), there is a column indicating whether each value is a presence point (1) or absence point (0).

Note that in this vignette, the validation data are read from .rds files because the original file was too big to be included in the package. However, the original file can be downloaded through the GUI or at https://github.com/smwoodman/eSDM-data

Use the function `sf::st_as_sf` to convert a data frame with lon/lat coordinates to an `sf` object.

```{r}
# Import and process validation data
# valid.data <- read.csv("eSDM_Validation_data_all.csv", stringsAsFactors = FALSE)
valid.data <- readRDS(system.file("extdata/eSDM_Validation_data_all.rds", package = "eSDM"))%>% 
  arrange(source, lat, lon) %>% 
  mutate(pres_abs = ifelse(pres_abs > 0, 1, 0)) %>% #For demonstration purposes; pres_abs column is already binary
  st_as_sf(coords = c("lon", "lat"), crs = 4326, agr = "constant") %>%
  st_transform(st_crs(base.geom))

# Extract the line transect and home range validation data
valid.data.lt <- valid.data %>% filter(source == "Becker_et_al_2016")
valid.data.hr <- valid.data %>% filter(source == "Irvine_et_al_2014")

# Summarize the number of presence and absence points
valid.data %>% 
  st_set_geometry(NULL) %>% 
  group_by(source) %>%  
  summarize(pres = sum(pres_abs == 1), 
            abs = sum(pres_abs == 0)) %>% 
  knitr::kable(caption = "Validation data summary")
```

Now we can calculate the AUC and TSS metrics for the original and overlaid predictions. The displayed table is from Table 3 of Woodman _et al._ (2019). Note that `evaluation_metrics` requires that validation data have the same coordinate reference system as the predictions being evaluated.

```{r, eval=FALSE}
# Calculate evaluation metrics with different validation data sets; code not run
names.1 <- c(
  "Model_B_orig", "Model_H_orig", "Model_R_orig", 
  "Model_B_overlaid", "Model_H_overlaid", "Model_R_overlaid"
)

eval.lt <- data.frame(do.call(rbind, list(
  eSDM::evaluation_metrics(model.b.sf, 1, st_transform(valid.data.lt, 4326), "pres_abs"), 
  eSDM::evaluation_metrics(model.h.sf, 1, st_transform(valid.data.lt, 4326), "pres_abs"), 
  eSDM::evaluation_metrics(model.r.sf, 1, valid.data.lt, "pres_abs"), 
  eSDM::evaluation_metrics(over1.sf, 1, valid.data.lt, "pres_abs"), 
  eSDM::evaluation_metrics(over2.sf, 1, valid.data.lt, "pres_abs"), 
  eSDM::evaluation_metrics(over3.sf, 1, valid.data.lt, "pres_abs")
))) %>%
  mutate(Preds = names.1) %>%
  dplyr::select(Preds, AUC_LT = X1, TSS_LT = X2)

eval.hr <- data.frame(do.call(rbind, list(
  eSDM::evaluation_metrics(model.b.sf, 1, st_transform(valid.data.hr, 4326), "pres_abs"), 
  eSDM::evaluation_metrics(model.h.sf, 1, st_transform(valid.data.hr, 4326), "pres_abs"), 
  eSDM::evaluation_metrics(model.r.sf, 1, valid.data.hr, "pres_abs"), 
  eSDM::evaluation_metrics(over1.sf, 1, valid.data.hr, "pres_abs"), 
  eSDM::evaluation_metrics(over2.sf, 1, valid.data.hr, "pres_abs"), 
  eSDM::evaluation_metrics(over3.sf, 1, valid.data.hr, "pres_abs")
))) %>%
  mutate(Preds = names.1) %>%
  dplyr::select(Preds, AUC_HR = X1, TSS_HR = X2)

eval.combo <- data.frame(do.call(rbind, list(
  eSDM::evaluation_metrics(model.b.sf, 1, st_transform(valid.data, 4326), "pres_abs"), 
  eSDM::evaluation_metrics(model.h.sf, 1, st_transform(valid.data, 4326), "pres_abs"), 
  eSDM::evaluation_metrics(model.r.sf, 1, valid.data, "pres_abs"), 
  eSDM::evaluation_metrics(over1.sf, 1, valid.data, "pres_abs"), 
  eSDM::evaluation_metrics(over2.sf, 1, valid.data, "pres_abs"), 
  eSDM::evaluation_metrics(over3.sf, 1, valid.data, "pres_abs")
))) %>%
  mutate(Preds = names.1) %>%
  dplyr::select(Preds, AUC = X1, TSS = X2)
```

```{r}
read.csv(system.file("extdata/Table3.csv", package = "eSDM")) %>%
  filter(grepl("Model_", Predictions)) %>% 
  dplyr::select(Predictions, AUC, TSS, `AUC-LT` = AUC.LT, `TSS-LT` = TSS.LT, 
                `AUC-HR` = AUC.HR, `TSS-HR` = TSS.HR) %>% 
  knitr::kable(caption = "Evaluation metrics", digits = 3, align = "lcccccc")
```

We can see that for each set of predictions, the original and overlaid evaluation metrics are quite similar, again showing that the overlay conserved the predicted blue whale distributions.


## Create and evaluate ensemble predictions

Before creating the ensembles, we must rescale the overlaid predictions. We rescaled the predictions using the abundance rescaling method and an abundance of 1648 (Becker _et al._ 2016). Using the sum to 1 rescaling method would result in ensembles with similar distribution patterns, but the actual density values could not be used to provide a meaningful abundance estimate.

`ensemble_rescale` requires a single `sf` object that contains all of the data being rescaled. Thus, we extract the prediction and variance values before using the rescaling function.


```{r}
# Rescale predictions
over.sf <- bind_cols(
  over1.sf %>% st_set_geometry(NULL) %>% dplyr::select(pred_bm1 = pred_bm, var1 = variance), 
  over2.sf %>% st_set_geometry(NULL) %>% dplyr::select(pred_bm2 = pred_bm, var2 = variance), 
  over3.sf %>% st_set_geometry(NULL) %>% dplyr::select(pred_bm3 = pred_bm, var3 = variance)
) %>% 
  st_sf(geometry = base.geom, agr = "constant")

over.sf.rescaled <- ensemble_rescale(
  over.sf, c("pred_bm1", "pred_bm2", "pred_bm3"), "abundance", 1648, 
  x.var.idx = c("var1", "var2", "var3")
)

# Check that overlaid predictions predict expected abundance
eSDM::model_abundance(over.sf.rescaled, "pred_bm1")
eSDM::model_abundance(over.sf.rescaled, "pred_bm2")
eSDM::model_abundance(over.sf.rescaled, "pred_bm3")

summary(over.sf.rescaled)
```

We can see that the prediction values are now much more comparable, and thus a subset of the predictions will not contribute disproportionately to an ensemble.

We also must calculate the ensemble weights, which must be manually created. For the example analysis, we use several weighting methods: equal weights (i.e., unweighted), AUC-based weights, TSS-based weights, and weights calculated as the inverse of the prediction variance. Each set of weights must sum to 1, or each row must sum to 1 in the case of the weights calculated as the inverse of the prediction variance. Note that when calculating evaluation metrics, a message is printed if any of the validation data points do not intersect with a prediction polygon.

```{r}
# Calculate ensemble weights
e.weights <- list(
  eSDM::evaluation_metrics(over1.sf, 1, valid.data, "pres_abs"), 
  eSDM::evaluation_metrics(over2.sf, 1, valid.data, "pres_abs"), 
  eSDM::evaluation_metrics(over3.sf, 1, valid.data, "pres_abs")
)

over.df.resc.var <- over.sf.rescaled %>% 
  dplyr::select(var1, var2, var3) %>% 
  st_set_geometry(NULL)

e.weights.unw <- c(1, 1, 1) / 3
e.weights.auc <- sapply(e.weights, function(i) i[1]) / sum(sapply(e.weights, function(i) i[1]))
e.weights.tss <- sapply(e.weights, function(i) i[2]) / sum(sapply(e.weights, function(i) i[2]))
e.weights.var <- data.frame(t(apply(
  1 / over.df.resc.var, 1, function(i) {i / sum(i, na.rm = TRUE)}
)))

e.weights.unw
e.weights.auc
e.weights.tss
head(e.weights.var)
```

Finally, we can create the ensembles. We calculate the ensemble uncertainty values with the among-model variance.

```{r}
### Create ensembles

# Unweighted; calculate CV because it is used in Fig. 4 plot
ens.sf.unw <- eSDM::ensemble_create(
  over.sf.rescaled, c("pred_bm1", "pred_bm2", "pred_bm3"),  w = e.weights.unw, 
  x.var.idx = NULL
) %>% 
  mutate(SE = sqrt(Var_ens), CV = SE / Pred_ens) %>% 
  dplyr::select(Pred_ens, SE, CV) %>% 
  st_set_agr("constant")

# Weights based on AUC
ens.sf.wauc <- eSDM::ensemble_create(
  over.sf.rescaled, c("pred_bm1", "pred_bm2", "pred_bm3"),  w = e.weights.auc, 
  x.var.idx = NULL
) %>% 
  mutate(SE = sqrt(Var_ens)) %>% 
  dplyr::select(Pred_ens, SE) %>% 
  st_set_agr("constant")

# Weights based on TSS
ens.sf.wtss <- eSDM::ensemble_create(
  over.sf.rescaled, c("pred_bm1", "pred_bm2", "pred_bm3"),  w = e.weights.tss, 
  x.var.idx = NULL
) %>% 
  mutate(SE = sqrt(Var_ens)) %>% 
  dplyr::select(Pred_ens, SE) %>% 
  st_set_agr("constant")

# Weights based on the inverse of the variance
ens.sf.wvar <- eSDM::ensemble_create(
  over.sf.rescaled, c("pred_bm1", "pred_bm2", "pred_bm3"),  w = e.weights.var, 
  x.var.idx = NULL
) %>% 
  mutate(SE = sqrt(Var_ens)) %>% 
  dplyr::select(Pred_ens, SE) %>% 
  st_set_agr("constant")
```

We could also have estimated the within-model ensemble uncertainty by using the `x.var.idx` argument, as demonstrated in the following code (not run).

```{r, eval=FALSE}
# Create an ensemble and calculate within-model uncertainty; code not run
ens.sf.unw.wmv <- eSDM::ensemble_create(
  over.sf.rescaled, c("pred_bm1", "pred_bm2", "pred_bm3"),  w = e.weights.unw,
  x.var.idx = c(var1, var2, var3)
) %>%
  mutate(SE = sqrt(Var_ens)) %>%
  dplyr::select(Pred_ens , SE)
```

Now we can calculate AUC and TSS scores for the ensembles and compare them to those of original and ensemble predictions. Again, the evaluation code is not run; the displayed table is Table 3 from Woodman _et al._ (2019).

```{r, eval=FALSE}
# Calculate evaluation metrics for ensembles; code not run
names.2 <- c(
  "Ensemble – unweighted", "Ensemble – AUC-based weights",
  "Ensemble – TSS-based weights", "Ensemble – variance-based weights"
)

eval.lt.ens <- data.frame(do.call(rbind, list(
  eSDM::evaluation_metrics(ens.sf.unw,  "Pred_ens", valid.data.lt, "pres_abs"), 
  eSDM::evaluation_metrics(ens.sf.wauc, "Pred_ens", valid.data.lt, "pres_abs"), 
  eSDM::evaluation_metrics(ens.sf.wtss, "Pred_ens", valid.data.lt, "pres_abs"), 
  eSDM::evaluation_metrics(ens.sf.wvar, "Pred_ens", valid.data.lt, "pres_abs")
))) %>%
  mutate(Preds = names.2) %>%
  dplyr::select(Preds, AUC_LT = X1, TSS_LT = X2)

eval.hr.ens <- data.frame(do.call(rbind, list(
  eSDM::evaluation_metrics(ens.sf.unw,  "Pred_ens", valid.data.hr, "pres_abs"), 
  eSDM::evaluation_metrics(ens.sf.wauc, "Pred_ens", valid.data.hr, "pres_abs"), 
  eSDM::evaluation_metrics(ens.sf.wtss, "Pred_ens", valid.data.hr, "pres_abs"), 
  eSDM::evaluation_metrics(ens.sf.wvar, "Pred_ens", valid.data.hr, "pres_abs")
))) %>%
  mutate(Preds = names.2) %>%
  dplyr::select(Preds, AUC_HR = X1, TSS_HR = X2) 

eval.combo.ens <- data.frame(do.call(rbind, list(
  eSDM::evaluation_metrics(ens.sf.unw,  "Pred_ens", valid.data, "pres_abs"), 
  eSDM::evaluation_metrics(ens.sf.wauc, "Pred_ens", valid.data, "pres_abs"), 
  eSDM::evaluation_metrics(ens.sf.wtss, "Pred_ens", valid.data, "pres_abs"), 
  eSDM::evaluation_metrics(ens.sf.wvar, "Pred_ens", valid.data, "pres_abs")
))) %>%
  mutate(Preds = names.2) %>%
  dplyr::select(Preds, AUC = X1, TSS = X2)
```

```{r}
read.csv(system.file("extdata/Table3.csv", package = "eSDM")) %>%
  dplyr::select(Predictions, AUC, TSS, `AUC-LT` = AUC.LT, `TSS-LT` = TSS.LT, 
                `AUC-HR` = AUC.HR, `TSS-HR` = TSS.HR) %>% 
  knitr::kable(caption = "Evaluation metrics", digits = 3, align = "lcccccc")
```

We can see that the ensemble with weights based on TSS values had the highest scores of the ensemble predictions, and mostly higher scores that the original predictions. We can visualize this ensemble to compare it with known blue whale habitat.

```{r, fig.width=7, fig.height=3}
# Simple code to visualize ensemble created with weights based on TSS values
plot_sf_3panel(
  rename(ens.sf.wtss, se = SE), "Pred_ens", main.txt = "Ensemble-TSS - ", 
  map.base = map.world, x.axis.at = c(-130, -125, -120)
)
```

Below is code to visualize the unweighted ensembles and this ensemble (i.e., create plots similar to Figs. 4 and 5 in Woodman _et al._). This code uses custom functions (located at `system.file("eSDM_vignette_helper.R", package = "eSDM")`) that leverage the `tmap` package to generate plots. However, by default this code is not run because each plot takes several minutes to generate.

```{r, fig.width=7, fig.height=4, eval=FALSE}
### Figure 4; code not run
library(tmap)

# Values passed to tmap_sdm - range of map
range.poly <- st_sfc(
  st_polygon(list(matrix(
    c(-132, -132, -116, -116, -132, 29.5, 49, 49, 29.5, 29.5), ncol = 2
  ))),
  crs = 4326
)
rpoly.mat <- matrix(st_bbox(range.poly), ncol = 2)

# Values passed to tmap_sdm - size of text labels and legend width
main.size <- 0.8
leg.size  <- 0.55
leg.width <- 0.43
grid.size <- 0.55

# Values passed to tmap_sdm - color scale info
blp1 <- tmap_sdm_help(ens.sf.unw, "Pred_ens")
blp2 <- tmap_sdm_help(ens.sf.unw, "CV")

# Plot of predictions (whales / km^-2)
tmap.obj1 <- tmap_sdm(
  ens.sf.unw, "Pred_ens", blp1, map.world, rpoly.mat, 
  "Unweighted ensemble - predictions", 
  main.size, leg.size, leg.width, grid.size
)
# Plot of SE values (with same color sceme as predictions)
tmap.obj2 <- tmap_sdm(
  ens.sf.unw, "SE", blp1, map.world, rpoly.mat, 
  "Unweighted ensemble - SE", 
  main.size, leg.size, leg.width, grid.size
)
# Plot of CV values
tmap.obj3 <- tmap_sdm(
  ens.sf.unw, "CV", blp2, map.world, rpoly.mat, 
  "Unweighted ensemble - CV", 
  main.size, leg.size, leg.width, grid.size
)

# Generate plot
tmap_arrange(
  list(tmap.obj1, tmap.obj2, tmap.obj3), ncol = 3, asp = NULL, outer.margins = 0.05
)
```

```{r, fig.height=9, fig.width=5.7, eval=FALSE}
### Figure 5; code not run

# Values passed to tmap_sdm - size of text labels and legend width
main.size <- 1.1
leg.size  <- 0.7
leg.width <- 0.6
grid.size <- 0.7

# Values passed to tmap_sdm - color scale info
blp1b <- tmap_sdm_help(ens.sf.wtss, "Pred_ens")
blp2b <- tmap_sdm_help_perc(ens.sf.wtss, "Pred_ens")

# Plot of predictions (whales / km^-2)
tmap.obj1 <- tmap_sdm(
  ens.sf.wtss, "Pred_ens", blp1, map.world, rpoly.mat, "Ensemble-TSS - Predictions", 
  main.size, leg.size, leg.width, grid.size
)
# Plot of SE values (with same color sceme as predictions)
tmap.obj2 <- tmap_sdm(
  ens.sf.wtss, "SE", blp1, map.world, rpoly.mat, "Ensemble-TSS - SE", 
  main.size, leg.size, leg.width, grid.size
)
# Plot of predictions (percentiles)
tmap.obj3 <- tmap_sdm(
  ens.sf.wtss, "Pred_ens", blp2b, map.world, rpoly.mat, "Ensemble-TSS - Predictions", 
  main.size, leg.size, leg.width, grid.size
)
# Plot of predictions (percentiles) with combined validation data presence points
tmap.obj4 <- tmap_sdm(
  ens.sf.wtss, "Pred_ens", blp2b, map.world, rpoly.mat, "Ensemble-TSS - Predictions", 
  main.size, leg.size, leg.width, grid.size
) + 
  tm_shape(filter(valid.data, pres_abs == 1)) + 
  tm_dots(col = "black", size = 0.04, shape = 19)

# Generate plot
tmap_arrange(
  list(tmap.obj1, tmap.obj2, tmap.obj3, tmap.obj4), ncol = 2, nrow = 2, 
  asp = NULL, outer.margins = 0.05
)
```


## References

Becker, E., Forney, K., Fiedler, P., Barlow, J., Chivers, S., Edwards, C., ... Redfern, J. (2016). Moving towards dynamic ocean management: how well do modeled ocean products predict species distributions? Remote Sensing, 8, 149. https://doi.org/10.3390/rs8020149

Hazen, E.L., Palacios, D.M., Forney, K.A., Howell, E.A., Becker, E., Hoover, A.L., ... Bailey, H. (2017). WhaleWatch: a dynamic management tool for predicting blue whale density in the California Current. Journal of Applied Ecology, 54, 1415-1428. https://doi.org/10.1111/1365-2664.12820

Irvine, L.M., Mate, B.R., Winsor, M.H., Palacios, D.M., Bograd, S.J., Costa, D.P. & Bailey, H. (2014). Spatial and temporal occurrence of blue whales off the U.S. West Coast, with implications for management. PLoS One, 9, e102959. https://doi.org/10.1371/journal.pone.0109485

Redfern, J.V., Moore, T.J., Fiedler, P.C., de Vos, A., Brownell, R.L., Forney, K.A., ... Heikkinen, R. (2017). Predicting cetacean distributions in data-poor marine ecosystems. Diversity and Distributions, 23, 394-408. https://doi.org/10.1111/ddi.12537

Woodman, S.M., Forney, K.A., Becker, E.A., DeAngelis, M.L., Hazen, E.L., Palacios, D.M., Redfern, J.V. (2019). eSDM: A tool for creating and exploring ensembles of predictions from species distribution and abundance models. Methods in Ecology and Evolution. 2019;10:1923-1933. https://doi.org/10.1111/2041-210X.13283
